How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-06-26 research/paper/draft.tex
added alternatives, completed assumptions

To transform the interim posteriors we have into the likelihoods we need, we 
start with the vacuously true statement
\begin{align}
\begin{split}
\label{eq:unity}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
\end{split}
\end{align}
We expand the last term using Bayes' Rule to get
\begin{align}
\begin{split}
\label{eq:bayes}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i}, 
\vec{\phi}^{*})] + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and 
$\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ 
as follows:
\begin{align}
\begin{split}
\label{eq:split}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{\phi}^{*})] - \log[p(\vec{d}_{i} | z_{i})] - 
\log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Canceling terms, we find
\begin{align}
\label{eq:cancel}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - 
\log[p(z_{i} | \vec{\phi}^{*})].
\end{align}

@@

This framework entails a number of choices and assumptions that must be 
addressed explicitly:
\begin{enumerate}
	\item The chosen functional form with parameters $\vec{\phi}$ must be 
capable of describing the true redshift distribution.
	\item The photometric data and redshift for each galaxy is independent 
of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior shared among all galaxies in the 
survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
	\item We must choose a hyperprior probability distribution 
$p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the 
probabilistic graphical model presented here is not the only one that could be 
proposed.
\end{enumerate}

@@

It is useful to translate some popular existing methods for deriving $n(z)$ 
from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.

@@

\begin{enumerate}
	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
samples of a single format, that of the piecewise constant parametrization, 
also referred to as a binned histogram parametrization and a sum of top hat 
functions.
\end{enumerate}

@@

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
function with known parameters $\vec{\theta}$.
	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
$N$ true redshifts $z_{i}$.
	\item \label{itm:scatter} Choose a true intrinsic scatter $\sigma$, and 
sample Gaussians $p(z'_{i} | z_{i}, \sigma) = \mathcal{N}(z_{i}, \sigma)$ to 
get "observed" redshifts $z'_{i}$ defining Gaussian likelihoods $p(\vec{d}_{i} 
| z'_{i}, \sigma) = \mathcal{N}(z'_{i}, \sigma)$.
	\item \label{itm:interim} Choose a parametrization and the parameters 
$\vec{\phi}^{*}$ of the interim prior.
	\item \label{itm:posterior} Create interim posteriors $p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*}) = p(\vec{d}_{i} | z_{i})\ p(z_{i} | 
\vec{\phi}^{*})$ in this parametrization.
\end{enumerate}

In all of the following validation tests, we use $10$ bins and $N = 10,000$ 
galaxies.  In Secs. \ref{sec:truth} and \ref{sec:likelihoods}, we use a flat 
interim prior.  This method for deriving mock data is referred to as the 
fiducial case, and variations on it will refer directly to the steps that are 
altered.

@@

>>>>

To transform the interim posteriors we have into the likelihoods we need, we 
start with the vacuously true statement
\begin{align}
\begin{split}
\label{eq:unity}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})].
\end{split}
\end{align}
We expand the last term using Bayes' Rule to get
\begin{align}
\begin{split}
\label{eq:bayes}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Fig. \ref{fig:pgm} has no direct connection between $\vec{d}_{i}$ and 
$\vec{\phi}$, so we can break up $\log[p(\vec{d}_{i} | z_{i}, \vec{\phi}^{*})]$ 
as follows:
\begin{align}
\begin{split}
\label{eq:split}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(\vec{d}_{i} | z_{i})] + \log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})]\\
& - \log[p(\vec{d}_{i} | z_{i})] - \log[p(\vec{d}_{i} | \vec{\phi}^{*})]\\
& + \log[p(\vec{d}_{i} | \vec{\phi}^{*})] - \log[p(z_{i} | \vec{\phi}^{*})].
\end{split}
\end{align}
Canceling terms, we find
\begin{align}
\label{eq:cancel}
\log[p(\vec{d}_{i} | z_{i})] =& \log[p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})] - 
\log[p(z_{i} | \vec{\phi}^{*})].
\end{align}

@@

This framework entails a number of choices and assumptions that must be 
addressed explicitly:
\begin{enumerate}
	\item The chosen functional form of $n(z)$ with parameters $\vec{\phi}$ 
must be capable of describing the true redshift distribution.
	\item The photometric data and redshift for each galaxy must be 
independent of the photometric data and redshifts of all other galaxies.
	\item There is one interim prior shared among all galaxies in the 
survey, and its parameters with known parameters $\vec{\phi}^{*}$ are known.
	\item We must choose a hyperprior probability distribution 
$p(\vec{\phi})$ over the hyperparameters.
	\item While we advocate for the approach of hierarchical inference, the 
probabilistic graphical model presented here is not the only one that could be 
proposed.
\end{enumerate}

@@

It is useful to translate some popular existing methods for deriving $n(z)$ 
from photo-$z$ PDFs into the mathematical framework of Sec. \ref{sec:model}.  
We do not advocate for these methods for reasons addressed in the Appendix.

\subsubsection{Stacking}
\label{sec:stacking}

The most common way to combine photo-$z$ PDFs into an estimator 
$\hat{n}_{stack}(z)$ for $n(z)$ is to stack them, which corresponds to
\begin{align}
\label{eq:stack}
\hat{n}^{stack}(z) &= \frac{1}{N}\ \sum_{i}^{N} p(z_{i} | \vec{d}_{i}, 
\vec{\phi}^{*}).
\end{align}

\subsubsection{Point Estimation}
\label{sec:point_estimation}

In some cases, estimators $\hat{n}(z)$ are obtained by assuming the photo-$z$ 
PDFs are effectively delta functions by reducing them to point estimates 
according to
\begin{align}
\label{eq:delta}
p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*}) &\approx \delta(z, \hat{z}_{i}^{point})
\end{align}
before applying Eq. \ref{eq:stack}.  The most common redshift point estimators 
are the mean $\hat{z}^{mean}_{i}$, median $\hat{z}^{median}_{i}$, and mode 
$\hat{z}^{mode}_{i}$ of the original photo-$z$ interim posterior.

@@

\begin{enumerate}
	\item \chippr is only applicable if the redshift interim posteriors and 
interim prior are accurate.
	\item \chippr currently only accepts photo-$z$ PDFs and produces $n(z)$ 
samples of a single format, that of the piecewise constant parametrization, 
also referred to as a binned histogram parametrization and a sum of top hat 
functions.
\end{enumerate}

@@

\begin{enumerate}
	\item \label{itm:true} Choose a true $n(z)$ that is a continuous 
function with known parameters $\vec{\phi}'$.
	\item \label{itm:sample} Sample the true $n(z)$ to create a catalog of 
$N$ true redshifts $z'_{i}$.
	\item \label{itm:scatter} Choose a true intrinsic scatter parameter 
$\sigma$, and sample Gaussians $p(z''_{i} | z'_{i}, \sigma) = 
\mathcal{N}(z'_{i}, \sigma)$ to get "observed" redshifts $z'_{i}$ defining 
Gaussian likelihoods $p(\vec{d}_{i} | z''_{i}, \sigma) = \mathcal{N}(z''_{i}, 
\sigma)$.
	\item \label{itm:interim} Choose a parametrization and the parameters 
$\vec{\phi}^{*}$ of the interim prior $p(z_{i} | \vec{\phi}^{*})$.
	\item \label{itm:posterior} Create interim posteriors $\log[p(z_{i} | 
\vec{d}_{i}, \vec{\phi}^{*})] = \log[p(\vec{d}_{i} | z'_{i})]\ \log[p(z_{i} | 
\vec{\phi}^{*})]$ in this parametrization.
\end{enumerate}

In all of the following validation tests, we use a piecewise constant 
parametrization in log-space with $10$ bins.  $N = 10,000$ galaxies.  In Secs. 
\ref{sec:truth} and \ref{sec:likelihoods}, we use a flat interim prior.  This 
method for deriving mock data is referred to as the fiducial case, and 
variations on it will refer directly to the steps that are altered.

@@

\appendix{}
